\section{Algorithms}
\label{app:algo}

Image generation, described by \Cref{algo:air_decoding}, is exactly the same for \gls{SQAIR} and \gls{AIR}.
\Cref{algo:sqair_prop,algo:sqair_disc} describe inference in \gls{SQAIR}. Note that \gls{disc} is equivalent to \gls{AIR} if no latent variables are present in the inputs.

If a function has multiple inputs and if not stated otherwise, all the inputs are concatenated and linearly projected into some fixed-dimensional space,\eg\Cref{l:rel_rnn,l:q_pres} in \Cref{algo:sqair_prop}.
Spatial Transformer ($\STN$,\eg \Cref{l:prop_glimpse} in \Cref{algo:sqair_prop}) has no learnable parameters: it samples a uniform grid of points from an image $\bx$, where the grid is transformed according to parameters $\bz^\mathrm{where}$.
$\operatorname{f_\phi^1}$~is implemented as a perceptron with a single hidden layer.
Statistics of $\qp$ and $\qd$ are a result of applying a two-layer \gls{MLP} to their respective conditioning sets.
Different distributions $\q$ do not share parameters of their \glspl{MLP}.
The \textit{glimpse encoder} $\operatorname{h_\phi^\mathrm{glimpse}}$ (\Cref{l:glimpse_enc_prop,l:glimpse_enc_final} in \Cref{algo:sqair_prop} and \Cref{l:glimpse_enc_disc} in \Cref{algo:sqair_disc}; they share parameters) and the \textit{image encoder} $\operatorname{h_\phi^\mathrm{enc}}$ (\Cref{l:img_encoder} in \Cref{algo:sqair_disc}) are implemented as two-layer \glspl{MLP} or \glspl{CNN}, depending on the experiment (see \Cref{app:mnist_details,app:duke_details} for details).

One of the important details of \gls{prop} is the proposal glimpse extracted in lines \Cref{l:prop_where,l:prop_glimpse} of \Cref{algo:sqair_prop}. 
It has a dual purpose. 
Firstly, it acts as an information bottleneck in \gls{prop}, limiting the flow of information from the current observation $\bxt$ to the updated latent variables $\bzt$.
Secondly, even though the information is limited, it can still provide a high-resolution view of the object corresponding to the currently updated latent variable, \textit{given} that the location of the proposal glimpse correctly predicts motion of this object.
Initially, our implementation used encoding of the raw observation ($\operatorname{h_\phi^\mathrm{enc}} \left( \bxt \right)$, similarly to \Cref{l:img_encoder} in \Cref{algo:sqair_disc}) as an input to the relation-\gls{RNN} (\Cref{l:rel_rnn} in \Cref{algo:sqair_prop}). We have also experimented with other bottlenecks: (1) low resolution image as an input to the image encoder and (2) a low-dimensional projection of the image encoding before the relation-\gls{RNN}. 
Both approaches have led to \textit{ID swaps}, where the order of explaining objects were sometimes swapped for different frames of the sequence (see \Cref{fig:id_swap} in \Cref{app:fail} for an example). 
Using encoded proposal glimpse extracted from a predicted location has solved this issue.
% \ak{should we create a "discussion" appendix and move this paragraph there?}

To condition \gls{disc} on propagated latent variables (\Cref{l:latent_enc} in \Cref{algo:sqair_disc}), we encode the latter by using a two-layer \gls{MLP} similarly to \cite{Zaheer2017},
\begin{equation}
    \blt = \sum_{i \in \mathcal{P}_t} \operatorname{MLP} \left( \bzt^{\mathrm{what}, i}, \bzt^{\mathrm{where}, i} \right).
\end{equation}
Note that other encoding schemes are possible, though we have experimented only with this one.

\begin{algorithm}
    \caption{Image Generation}
    \label{algo:air_decoding}
    \DontPrintSemicolon
    \SetKwInOut{Input}{Input}
    \SetKwInOut{Output}{Output}
    \SetSideCommentLeft
    \Input{
        $\bzt^\mathrm{what},
         \bzt^\mathrm{where}$ 
         - latent variables from the current time-step.\\
    }
    % 
    $\mathcal{O}_t = \operatorname{indices} \left( \bzt^\mathrm{what} \right)$ \tcp*{Indices of all present latent variables.}
    $\byt^0 = \bm{0}$\\
    \For{$i \in \mathcal{O}_t$}{
        $\byt^{\mathrm{att}, i} = \operatorname{f_\theta^\mathrm{dec}} \left( 
            \bzt^{\mathrm{what}, i} 
        \right)$ \tcp*{Decode the glimpse.}
% 
        $\byt^i = \byt^{i-1} + \STN^{-1} \left( 
            \byt^{\mathrm{att}, i}, \bzt^{\mathrm{where}, i} 
        \right)$
    }
    $\hat{\bx}_t \sim \mathcal{N} \left(\bm{x} \mid \bm{y}_n, \sigma^2_x \bm{I} \right)$\\
    \Output{$\hat{\bm{x}}$}
\end{algorithm}

\begin{algorithm}
    \caption{Inference for Propagation}
    \label{algo:sqair_prop}
    \DontPrintSemicolon
    \SetKwInOut{Input}{Input}
    \SetKwInOut{Output}{Output}
    \SetSideCommentLeft
    \Input{$\bm{x_t}$ - image at the current time-step,\\
         $\bz^\mathrm{what}_{t-1},
         \bz^\mathrm{where}_{t-1},
         \bz^\mathrm{pres}_{t-1}$ 
         - latent variables from the previous time-step\\
         $\bm{h}^T_{t-1}$ - hidden states from the previous time-step.
    }
    $\hR{t}{0}, \bm{z}^{\mathrm{what}, 0}_t, \bm{z}^{\mathrm{where}, 0}_t = \operatorname{initialize}()$\\
    $j = 0$ \tcp*{Index of the object processed in the last iteration.}
    \For{$i \in {\mathcal{O}_{t-1}}$}{
% 
        \uIf{$z_{t-1}^{\mathrm{pres}, i} == 0$}{
            \Continue
        }
%       
        $\hat{\bz}_t^{\mathrm{where}, i} = \operatorname{f_\phi^1} \left( \bz_{t-1}^{\mathrm{where}, i},
            \hT{t}{i} \right)$ \label{l:prop_where}\tcp*{Proposal location.}
% 
        $\hat{\bg}^i_t = \STN \left(
            \bxt, \hat{\bz}_t^{\mathrm{where}, i}
        \right)$ \label{l:prop_glimpse}\tcp*{Extract a glimpse from a proposal location.}
% 
    $\hat{\be}_t^i = \operatorname{h_\phi^\mathrm{glimpse}} \left( \hat{\bg}^i_t \right)$ \label{l:glimpse_enc_prop}\tcp*{Encode the proposal glimpse.}
% 
        $\textcolor{orange}{\bwt^{R, i}}, \hR{t}{i} = \Rr \left(
            \hat{\be}_t^i,
            \bz_{t-1}^{\mathrm{what}, i}, \bz_{t-1}^{\mathrm{where}, i}, 
            \hT{t-1}{i}, \hR{t}{j},
            \bzt^{\mathrm{what}, j}, \bzt^{\mathrm{where}, j}
        \right)$  
        \label{l:rel_rnn}\tcp*{Relational state, see \Cref{eq:propagation_state}.}
%         
        $\bzt^{\mathrm{where}, i} \sim \qp{\bm{z}^\mathrm{where}}{\bm{z}^{\mathrm{where}, k}_{t-1}, \textcolor{orange}{\bwt^{R, i}}}{\phi}$\\
%        
        $\bgt^i = \STN \left( 
            \bxt,
            \bzt^{\mathrm{where}, i}
        \right)$ \tcp*{Extract the final glimpse.}
    % 
    $\bet^i = \operatorname{h_\phi^\mathrm{glimpse}} \left( \bgt^i \right)$ \label{l:glimpse_enc_final}\tcp*{Encode the final glimpse.}
    % 
        $\textcolor{blue}{\bwt^{T, i}}, \hT{t}{i} = \RT \left(
            \bet^i,
            \bzt^{\mathrm{where}, i}, 
            \hT{t-1}{i}, \hR{t}{i}
        \right)$
        \label{l:time_rnn}\tcp*{Temporal state, see \Cref{eq:temporal_state}.}
%        
        $\bzt^{\mathrm{what}, i} \sim \qp {\bz^\mathrm{what}}{\bet^i, \bz_{t-1}^{\mathrm{what}, i}, \textcolor{orange}{\bwt^{R, i}}, \textcolor{blue}{\bwt^{T, i}}}{\phi}$\\
%        
        $\zt^{\mathrm{pres}, i} \sim \qp{z^\mathrm{pres}}{
            z_{t-1}^{\mathrm{pres}, i},
            \bzt^{\mathrm{what}, i},
            \bzt^{\mathrm{where}, i},
            \textcolor{orange}{\bwt^{R, i}},
            \textcolor{blue}{\bwt^{T, i}}}{\phi}$ \label{l:q_pres}\tcp*{\Cref{eq:q_prop_presence}.}
        $j = i$
    }
    \Output{
        $\bzt^{\mathrm{what}, \mathcal{P}_t},
        \bzt^{\mathrm{where}, \mathcal{P}_t},
        \bzt^{\mathrm{pres}, \mathcal{P}_t}$
    }
\end{algorithm}

\begin{algorithm}
    \caption{Inference for Discovery}
    \label{algo:sqair_disc}
    \DontPrintSemicolon
    \SetKwInOut{Input}{Input}
    \SetKwInOut{Output}{Output}
    \SetSideCommentLeft
    \Input{
        $\bm{x_t}$ - image at the current time-step,\\
        $\bzt^{\mathcal{P}_t}$ - propagated latent variables for the current time-step,\\
        $N$ - maximum number of inference steps for discovery.
    }
% 
    $\hD{t}{0}, \bm{z}^{\mathrm{what}, 0}_t, \bm{z}^{\mathrm{where}, 0}_t = \operatorname{initialize}()$\\
% 
    $j = \operatorname{max\_index} \left( \bzt^{\mathcal{P}_t} \right)$ \tcp*{Maximum index among the propagated latent variables.}
% 
    $\bet = \operatorname{h_\phi^\mathrm{enc}} \left( \bxt \right)$ \label{l:img_encoder}\tcp*{Encode the image.}
% 
    $\blt = \operatorname{h_\phi^\mathrm{enc}} \left( 
        \bzt^\mathrm{what},
         \bzt^\mathrm{where},
         \bzt^\mathrm{pres} 
     \right)$ \label{l:latent_enc}\tcp*{Encode latent variables.}
%  
    \For{$i \in [j + 1, \dots, j + N]$}{
        $\textcolor{pink}{\bwt^{D, i}}, \hD{t}{i} = \RD \left(
            \bet,
            \blt,
            \bzt^{\mathrm{what}, i-1}, 
            \bzt^{\mathrm{where}, i-1}, 
            \hD{t}{i-1} 
        \right)$\\
% 
        $\zt^{\mathrm{pres}, i} \sim \qd{z^\mathrm{pres}}{ \textcolor{pink}{\bwt^{D, i}}}{\phi}$\\
% 
        \If{$z^{\mathrm{pres}, i} = 0$}{
            break
        }
% 
        $\bzt^{\mathrm{where}, i} \sim \qd{\bz^\mathrm{where}}{\textcolor{pink}{\bwt^{D, i}}}{\phi}$\\
% 
        $\bgt^i = \STN\left( 
            \bxt,
            \bzt^{\mathrm{where}, i} 
        \right)$\\
% 
        $\bet^i = \operatorname{h_\phi^\mathrm{glimpse}} \left( \bgt^i \right)$ \label{l:glimpse_enc_disc}\tcp*{Encode the glimpse.}
% 
        $\bzt^{\mathrm{what}, i} \sim \qd{\bzt^\mathrm{what}}{\bet^i}{\phi}$
    }
    \Output{
       $\bzt^{\mathrm{what}, \mathcal{D}_t},
        \bzt^{\mathrm{where}, \mathcal{D}_t},
        \bzt^{\mathrm{pres}, \mathcal{D}_t}$
    }
\end{algorithm}