\documentclass{article}

% \usepackage{corl_2019} % initial submission
\usepackage[final]{corl_2019} % Uncomment for the camera-ready ``final'' version

\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{times}
% numbers option provides compact numerical references in the text. 
\usepackage[numbers]{natbib}
\usepackage{multicol}
\usepackage{graphicx}
% \usepackage{subfigure}
%\usepackage{cleveref}
\usepackage[noabbrev,capitalize]{cleveref}
\usepackage{caption}
\usepackage{subcaption}
% \usepackage[skip=5pt,font=small]{caption}
% \usepackage[skip=5pt,font=small]{subcaption}
\usepackage{titlesec} % % section spacing


% for tables side-by-side
\usepackage{geometry}
\usepackage{array, caption, floatrow, tabularx, makecell, booktabs}%
\captionsetup{labelfont = sc}
\setcellgapes{3pt}

\usepackage[romanian,english]{babel} \usepackage{combelow} %for stefan's name

\usepackage{mlmacros}

\usepackage[T1]{fontenc}
\usepackage[acronym,smallcaps,nowarn,section,nogroupskip,nonumberlist]{glossaries}
\renewcommand*{\glstextformat}[1]{\textcolor{black}{#1}}
\glsdisablehyper % disable hyperlinks to gls commands globally

\input{glossary}

\usepackage{todonotes} % this generates a lot of warnings

\newcommand{\ak}[1]{\textcolor{blue}{\textbf{ak}: #1}}
\newcommand{\AK}[1]{\todo{\ak{#1}}}

\newcommand{\opj}[1]{\textcolor{purple}{\textbf{opj}: #1}}
\newcommand{\OPJ}[1]{\todo{\opj{#1}}}


\renewcommand{\citep}[1]{(\citet{#1})}

\variables{a,g,t,x,y,z}


\title{End-to-end Recurrent Multi-Object Tracking and Trajectory Prediction with Relational Reasoning}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

% NOTE: authors will be visible only in the camera-ready (ie, when using the option 'final'). 
% 	For the initial submission the authors will be anonymized.


% \author{
%   Fabian Fuchs\\
%   Applied AI Lab\\
%   University of Oxford\\
%   \texttt{fabian@robots.ox.ac.uk} \\
%   \And
%   Adam R. Kosiorek\\
%   Applied AI Lab\\
%   University of Oxford\\
%   \texttt{adamk@robots.ox.ac.uk} \\
%   \And
%   Li Sun\\
%   Applied AI Lab\\
%   University of Oxford\\
%   \texttt{kevin@robots.ox.ac.uk} \\
%   \And
%   Oiwi Parker Jones\\
%   Applied AI Lab\\
%   University of Oxford\\
%   \texttt{oiwi@robots.ox.ac.uk} \\
%   \And
%   Ingmar Posner\\
%   Applied AI Lab\\
%   University of Oxford\\
%   \texttt{ingmar@robots.ox.ac.uk} \\
% }

\author{
 Fabian B. Fuchs,\, Adam R. Kosiorek,\, Li Sun,\, Oiwi Parker Jones,\,Ingmar Posner\\
 Applied AI Lab,  University of Oxford \\
 \texttt{\{fabian, adamk, kevin, oiwi, ingmar\}@robots.ox.ac.uk} \\
}

\begin{document}
\maketitle

%===============================================================================

\begin{abstract}
The majority of contemporary object-tracking approaches used in autonomous vehicles do not model interactions between objects. This contrasts with the fact that objects' paths are not independent: a cyclist might abruptly deviate from a previously planned trajectory in order to avoid colliding with a car.
Building upon \textsc{hart}, a neural, class-agnostic single-object tracker, we introduce a multi-object tracking method (\textsc{mohart}) capable of \emph{relational reasoning}. Importantly, the entire system, including the understanding of interactions and relations between objects, is class-agnostic and learned simultaneously in an end-to-end fashion.
We find that the addition of relational-reasoning capabilities to \textsc{hart} leads to consistent performance gains in tracking as well as future trajectory prediction on several real-world datasets (MOTChallenge, UA-DETRAC, and Stanford Drone dataset), particularly in the presence of ego-motion, occlusions, crowded scenes, and faulty sensor inputs.
Finally, based on controlled simulations, we propose that a comparison of \textsc{mohart} and \textsc{hart} may be used as a novel way to measure the degree to which the objects in a video depend upon each other as they move together through time. 


\end{abstract}


% Two or three meaningful keywords should be added here
\keywords{Multi-Object Tracking, Trajectory Prediction, Relational Reasoning} 

%===============================================================================

\input{1_introduction}
\input{1b_related_work}
\input{2_model}
\input{3_validation}
\input{4_experiments}
\input{5_discussion}


\section*{Acknowledgements}
We thank Stefan Saftescu for his contributions, particularly for integrating the Stanford Drone Dataset, and Adam Golinski as well as Stefan Saftescu for proof-reading. This research was funded by the EPSRC AIMS Centre for Doctoral Training at Oxford University and an EPSRC Programme Grant (EP/M019918/1).
We acknowledge use of Hartree Centre resources in this work. The STFC Hartree Centre is a research collaboratory in association with IBM providing High Performance Computing platforms funded by the UK's investment in e-Infrastructure. The Centre aims to develop and demonstrate next generation software, optimised to take advantage of the move towards exa-scale computing.
% no \bibliographystyle is required, since the corl style is automatically used.
\newpage\bibliography{library}  % .bib

\newpage\clearpage\input{6_appendix.tex}

\end{document}

