I would like to thank

\begin{itemize}
	\item My mum for encouraging my creativity and answering my endless questions when I was growing up.
	\item  My dad for pushing me to work hard, teaching me about business and negotiation, and mentoring me in my carrer so far. It was him who planted the idea of doing a PhD in my mind.
	\item My grandparents for raising me, always being there for me, providing a secure base, and showing me what it means to be loved.
	\item My dear childhood friend Rafał Worobiew for taking care of me at a crucial moment, and showing me an alternative path.
	\item Prof. Barbara Siemiątkowska of the Warsaw Univeristy of Technology for teaching an intro to computer vision course, which is where my journey to AI started.
	\item Andrzej Ruta, at the time a team lead at Samsung R\&D in Warsaw, who hired the clueless student that I was as an intern in his computer vision research team.
	\item Patrick van der Smagt of the Technical University of Munich (at the time) for teaching my first formal ML course. 
	\item Martin Engelcke who, during a chance encounter, convinced me to apply to Oxford and specifically to his supervisor Ingmar Posner's group.
	\item My supervisor Ingmar Posner who welcomed me to his group at Oxford, initiated me into the land of independent research, but was also able to tolerate my sometimes harsh remarks and moves, and ultimately allowed me to move away.
	\item My supervisor Yee Whye Teh who allowed me to join his group in my 2nd year of PhD.
	\item Danilo Rezende for hosting me as an intern at DeepMind.
	\item Geoff Hinton for inviting me for an internship after a brief discussion at a conference.
	\item My friends and colleagues who turned this PhD into an unforgettable journey: Alex Bewley, Julie Dequire, Neil Dhir, Martin Engelcke, Fabian Fuchs, Adam Goliński, Oliver Groth, Sandy Huang, Hyunjik Kim, Tuan Anh Le, Milan Muso, Tom Rainforth, Ștefan Săftescu, Hillary Shakespeare, Jimmy Shi, Georgi Tinchev, Markus Wulfmeier, and many others.
\end{itemize}

The funding for this thesis was provided by the University of Oxford and a scholarship funded by Google Deepmind. I extend my thanks to Ingmar Posner for helping secure the scholarship.